\documentclass{article}
\usepackage{indentfirst}

\begin{document}

\title{WRENCH Pedagogic Modules: Distributed Computing Courseware}
\author{Ryan Tanaka}

\maketitle

\section{Introduction}
The WRENCH Pedagogic Modules are a guided set of hands on, educational
activities aimed at teaching distributed computing concepts using the WRENCH
workflow management system simulation framework introduced in \cite{wrench}.
This simulation based, distributed computing courseware was developed with the
goal of accelerating distributed computing education in the classroom by
affording higher student participation over traditional methods and by providing
a more effective means of pedagogy. To achieve this goal, we utilize the WRENCH
simulation framework along with contemporary web technologies. These modules are
not meant to replace a conventional education on distributed computing, rather
they are meant to supplement it so as to deliver a modern, yet effective
educational  experience.

This paper is organized as follows. First I elaborate on the benefits of using
simulation for teaching  distributed computing. Second, the design and layout of
these modules are described. Next, I discuss results we have obtained from two
initial studies. Finally, I detail future plans to build upon the current set of
modules and also to submit our work to a high performance computing education
conference.

\section{Using Simulations}
Conventional distributed computing education often consists of theory
incorporated with a mix of hands-on learning experiences using real world
platforms. While this is a widely accepted method of teaching distributed computing,
it has been shown that both \textit{logistical challenges} and \textit{pedagogical challenges}
stifle its effectiveness \cite{smpi-courseware}. Logistical challenges arise when
students need access to cyber infrastructure platforms on which they can execute
hands-on experiments. Not all institutions have such access to various 
cyber infrastructure systems. Furthermore, access to these systems is limited due
to quantity and availability constraints. Assuming access to one type of system is
given, pedagogical challenges are still present. First, training or "on boarding" is necessary
to get students familiarized with that system. Second, students are limited to 
predetermined cyber infrastructure configurations, therefore limiting their ability to
experiment with "What if?" scenarios. For example, it will not be possible for students to 
run hands on experiments on a wide variety of cyber infrastructure scenarios without
access to multiple systems due to the aforementioned 
logistical challenges. 

With the use of WRENCH simulation software, arbitrary cyber infrastructure can be simulated
on a single commodity computer, therefore directly addressing both logistical and 
pedagogical challenges. Student participation only requires a desktop or laptop with modest
hardware specifications. Executing WRENCH simulations requires basic knowledge of how to
run programs from a command line and time spent "on boarding" in this case is
negligible. In regards to pedagogy, this simulation software affords students the ability
to experiment with "What if?" scenarios, thereby enriching their understanding of the
theoretical concepts at hand. The underlying simulation engine used in WRENCH is provided
by the SimGrid distributed systems simulation framework \cite{simgrid}, whose simulation models have been 
rigorously validated over the years. 

We are confident that with the use of such accurate simulation
software along with a web based, guided curriculum, a valuable contribution can be made to
distributed computing pedagogy. Next, I introduce the WRENCH Pedagogic Modules.


\section{The WRENCH Pedagogic Modules}
The WRENCH Pedagogic Modules are a guided set of hands on, educational activities focused
on teaching distributed computing concepts through the use of interactive simulations. It currently 
contains five guided lessons hosted online, four WRENCH simulations, and an interactive
browser-based dashboard from which users can run those simulations and view results through
a number of different custom visualizations. 

Each of the five lessons contain student learning objectives, background information on the specific
distributed computing concept being covered, and a handful of questions that lead students toward
achieving each learning objective in a moment of revelation. Many of these questions require students
to formulate a simple mathematical model based on some described scenario, then have them check that
model against results returned from running the simulation through the dashboard. Then the scenario
is changed so that students can be exposed to the "What if?" scenarios. For example, 
a series of performance related questions might be asked regarding a high performance computing
cluster given some algorithm and workload. First, students are asked to investigate this question
assuming the cluster has a single compute node. Then, follow up questions are asked that modify
the scenario such that the cluster now has multiple compute nodes. In both scenarios, students can
verify their results by running the simulations through our dashboard. 

In order to address the issue of dependencies, we host these modules online at
\underline{https://wrench-project.org/wrench-pedagogic-modules/} and containerize the dashboard
along with each of the simulations using Docker. As a result, students need only a modest computer, an internet
connection, a web browser, and Docker or Vagrant installed to participate. 

\section{Preliminary Results}


\section{Future Work}

\bibliographystyle{ieeetr}
\bibliography{sources}

\end{document}
